DRAFT \\ From physical principles to classical and quantum \\ Hamiltonian and Lagrangian particle mechanics

### 2016-11-01 2015CompleteDerivation/completeDerivation.tex
First pass at quantum

\begin{assump}[Infinitesimal reducibility]\label{ass:infinitesimal_reducibility}
	The system under study is composed of an irreducible homogeneous material and as a whole undergoes deterministic and reversible evolution.
\end{assump}

@@

\begin{prop}\label{unitary_evolution}
	A deterministic and reversible evolution map $\mathcal{T}_{\Delta t}: \Psi \rightarrow \Psi$ for an irreducible material is a unitary operator. That is: $\langle \mathcal{T}_{\Delta t} \psi_1, \mathcal{T}_{\Delta t} \psi_2 \rangle = \langle \psi_1 , \psi_2 \rangle$ where $\psi_1, \psi_2 \in \Psi$. The time evolution for the wave function $\Upsilon_q \mathcal{T}_{\Delta t} \Upsilon^q : S(\mathcal{Q}, dq) \rightarrow S(\mathcal{Q}, dq)$ is a $C^{\infty}$-diffeomorphism.
\end{prop}

\begin{justification}
\end{justification}

@@

\begin{justification}
\end{justification}

Note: we get time dependent Schroedinger. No basis for time independent.

>>>>

\begin{assump}[Infinitesimal reducibility]\label{ass:irreducibility}
	The system under study is composed of an irreducible homogeneous material and as a whole undergoes deterministic and reversible evolution.
\end{assump}

@@

\begin{prop}\label{prop:unitary_evolution}
	A deterministic and reversible evolution map $\mathcal{T}_{\Delta t}: \Psi \rightarrow \Psi$ for an irreducible material is a unitary operator. That is: $\langle \mathcal{T}_{\Delta t} \psi_1, \mathcal{T}_{\Delta t} \psi_2 \rangle = \langle \psi_1 , \psi_2 \rangle$ where $\psi_1, \psi_2 \in \Psi$.
\end{prop}

\begin{justification}
	We claim $\mathcal{T}_{\Delta t}: \Psi \rightarrow \Psi$ is a unitary operator. Let $\psi \in \Psi$ be a state. $|\psi|^2 = |\mathcal{T}_{\Delta t} \psi|^2$ because deterministic and reversible evolution does not change the amount of material. Let $\psi_1, \psi_2 \in \Psi$ be two states. $\mathcal{T}_{\Delta t}(a\psi_1+b\psi_2) = a \mathcal{T}_{\Delta t} \psi_1 + b \mathcal{T}_{\Delta t} \psi_2$ with $a,b \in \mathcal{C}$ as the composition of the evolution is the evolution of the composition: $\mathcal{T}_{\Delta t}$ is a linear operator. $\mathcal{T}_{\Delta t} \mathsf{P}_{\psi_1} (\psi_2) = \mathsf{P}_{\mathcal{T}_{\Delta t}\psi_1} (\mathcal{T}_{\Delta t}\psi_2)$ as the evolved part of $\psi_2$ that was prepared like $\psi_1$ is the part of the evolved $\psi_2$ that ends prepared like the evolved $\psi_1$. We have $\mathcal{T}_{\Delta t} \mathsf{P}_{\psi_1} (\psi_2) = \mathcal{T}_{\Delta t} \langle \psi_1 , \psi_2 \rangle \psi_1 = \langle \psi_1 , \psi_2 \rangle \mathcal{T}_{\Delta t} \psi_1 = \mathsf{P}_{\mathcal{T}_{\Delta t}\psi_1} (\mathcal{T}_{\Delta t}\psi_2) = \langle \mathcal{T}_{\Delta t}\psi_1 , \mathcal{T}_{\Delta t}\psi_2 \rangle \mathcal{T}_{\Delta t} \psi_1$. $\langle \psi_1 , \psi_2 \rangle = \langle \mathcal{T}_{\Delta t}\psi_1 , \mathcal{T}_{\Delta t}\psi_2 \rangle$. $\mathcal{T}_{\Delta t}$ is unitary.
\end{justification}

@@

\begin{justification}
	We claim the state evolves according to Schroedinger's equation. Let  $\mathcal{T}_{dt}: \Psi \rightarrow \Psi$ be the evolution for an infinitesimal time interval $dt$. As it is unitary, it can be expressed as $\mathcal{T}_{dt} = I + \frac{H dt}{\imath \hbar}$ where $H : \psi \rightarrow \psi$ is a self-adjoint operator. Let $\psi_t \in \Psi$ be a state and $\psi_{t+dt} \in \Psi$ its evolution after an infinitesimal time interval. We have $\mathcal{T}_{dt} \psi_t = \psi_{t+dt} = \psi_t + \frac{H dt}{\imath \hbar} \psi_t$. $\imath \hbar \frac{\psi_{t+dt} - \psi_t}{dt} = \imath \hbar \partial_t \psi_t = H\psi_t$.
\end{justification}

We recognize the Schroedinger equation, the time evolution equation for a quantum particle state.

Note, though, that nothing tells us that the system must be in an eigenfunction of $H$ or that it the ground state is somehow preferred. In fact, the eigenfunctions of $H$ may not be part of $S(\mathcal{Q}, q^i)$ and therefore may not be physical. In other words: we have no basis for the time independent Schroedinger equation. This is actually consistent. For example, for a system to reach its ground state it has to be able to radiate energy: it cannot isolated. Therefore this dynamic cannot be described under the deterministic and reversible assumption. This already hints that deterministic and reversible evolution is not the only dynamic possible. Other considerations outside of the scope of this work must be used to recover this behavior (e.g. thermodynamics or decoherence).

\subsection{Time dependence and kinematic assumption}

The next step would be to integrate the quantum description with time dependent evolution and the kinematic assumption. We will only provide a sketch, though, without going through the details. These would take a significant amount of space while not providing any significantly new insight. Also, given that we are limited to single particles (i.e. no fields) without spin (i.e. minimal material only described by position), our results are limited to finding the Klein-Gordon equation modified with electromagnetic interactions. While this is still of note, the equation is of limited physical use and therefore may not warrant the space necessary for a more rigorous derivation.

As for the classical case, the first thing to do is to extend $\mathcal{Q}$ to include time. This gives us the same $\mathcal{M}$ as before. A quantum state is a complex valued distribution $\psi : \mathcal{M}_{t=t_0} \rightarrow \mathbbm{C}$ over an hypersurface at constant time. An evolution $\bar{\psi} : \mathcal{M} \rightarrow \mathbbm{C}$ is a complex valued function defined on the whole $\mathcal{M}$. The operator for the quantity conjugate of time is $E=\imath\hbar\partial_t$.

Next we need to extend integration. In the quantum case, this means to make sure the inner product $\langle \psi | \phi \rangle$ remains unchanged under all variable changes, including time. This allows to describe the amounts of material and their distribution over fragments in a way that is independent of our description.

As in the classical case, we use $s$ as the evolution parameter. Deterministic and reversible evolution $\mathcal{T}_{ds}$ will be a unitary transformation defined over this extended inner product, as the amounts of material get conserved. Continuous evolution will admit an invariant Hamiltonian $\mathcal{H}$ such that $\imath \hbar \partial_s \psi = \mathcal{H} \psi$. As for the composite state evolution in the classical case, $\bar{\psi}$ remains unchanged by the evolution as it transports the distribution along $s$. Therefore $\mathcal{H} \hat{\psi} = \mathcal{h} \hat{\psi}$ is an eigenfunction of $\mathcal{H}$. As in the classical case, we use $\mathcal{H}$ to be zero to identify the state space: the temporal d.o.f. is not independent and therefore there is a constraint between the variables. Therefore we have $\mathcal{H} \psi = 0$. This is the time dependent evolution equation.

Note that states are indeed eigenfunctions of the invariant Hamiltonian $\mathcal{H}$: here we do have a justification, as opposed to the standard Hamiltonian $H$ case. For example, a massive particle may have fragments with different energy, but they all must have the same mass. Also note that $\mathcal{H} \psi = 0$ is basically the form of all equations (i.e. Klein-Gordon, Dirac, Maxwell, ...) in quantum field theory. The respective invariant Hamiltonian $\mathcal{H}$ is the operator associated with each equation.

Having handled the time dependent case, we can turn our attention to the kinematic equivalence assumption. While the trajectory of a fragment is not well defined in either phase space or physical space, what we can describe remains the same under the assumption. Therefore the distributions still need to be transported from state space to kinematic variables up to a constant factor. Which means we still have a linear relationship between velocity and conjugate momentum. In terms of the quantum operators: $m\, d_s Q^\alpha = g^{\alpha\beta} P_{\beta} - A^\alpha = m\frac{[Q^\alpha, \mathcal{H}]}{\imath \hbar}$.

This leads to the same general for the invariant Hamiltonian $\mathcal{H}=\frac{1}{2m}((P_\alpha-A_\alpha)g^{\alpha\beta}(P_\beta-A_\beta) + m^2 c^2) = \frac{1}{2m}((-\imath\hbar\partial_\alpha-A_\alpha)g^{\alpha\beta}(-\imath\hbar\partial_\beta-A_\beta) + m^2 c^2)$. In the free particle case, we have $\frac{1}{2m}(-\hbar^2 \partial^\alpha \partial_\alpha + m^2 c^2) \bar{\psi} = 0$ which is the Klein-Gordon equation. We can also write the operator for kinetic momentum as $mU^\alpha = P^\alpha - A^\alpha = i\hbar D^\alpha$ where $D^\alpha = \partial^\alpha - \frac{A^\alpha}{\imath \hbar}$ is the covariant gauge derivative. We have $(\frac{1}{2}mU^\alpha U_\alpha + \frac{1}{2} m c^2) \bar{\psi} = \frac{1}{2m}(-\hbar^2 D^\alpha D_\alpha + m^2 c^2) \bar{\psi} = 0$ which is the Klein-Gordon equation with electromagnetic interaction.

Note how the covariant derivative is, up to a constant, kinetic momentum. This motivates why it is such an important operator. As for the Klein-Gordon equation, the only thing it does is to impose that the norm of the four velocity is the speed of light, a condition much more trivial one may have anticipated.

It should be clear, even without the details, that all the assumptions come together even in the quantum case. They give more insight to all the pieces in a comprehensive way and help tear down the artificial walls between the different physical theories. This allows to form a picture of fundamental physics that is more unified, something that has been sorely missing.

\subsection{Non deterministic evolution and the projection postulate}

As we made it clear from the start, deterministic and reversible evolution is an assumption. And as we seen in the justifications for assumptions \ref{ass:infinitesimal_reducibility} and \ref{ass:irreducibility} we cannot have determinism at all levels of decomposition while transferring information from one system to another.

In this section we will see how non deterministic processing become necessary when describing measurements for irreducible systems. But we'll see that, under the same assumption of irreducibility, we are able to make a connection between the output of deterministic and non deterministic processes.

Suppose we have a quantum particle state $\psi$ we want to identify. Suppose that we have a detector that is able to capture the particle. If it's captured, we will know that the particle was in the region of space enclosed by the detector.

Suppose that we create an array of such detectors. This will allow to detect the position of the particle with a resolution given by the aperture of each detector.

Suppose now, though, that the quantum particle state is spread over a region significantly greater than a single detector but smaller than the array. What will it happen?

It is clear that it has to be absorbed as it has no way to escape. It is also clear that it must be absorbed whole: as it is an irreducible system, the detector cannot partially interact with it, it cannot half absorb it. So a single detector will capture it. It is also clear that it cannot always be the same detector: that corresponds to states where the particle is spread within the aperture of that detector. In other words: as the particle hits the array, one detector triggers randomly based on the distribution of the particle.

The process is non deterministic in the sense that the state of the particle is not enough to predict which detector will trigger. Maybe it depends on the state of the detector, maybe it depends on the unstated part of the particle. The point is that a single take does not tell us much about the incoming particle states.

We then repeat the process, making sure the incoming particle is in the same state. Given a single "pure" incoming particle state, we get a "mixed" statistical distribution describing the output of the array.\footnote{Note that this is different from the model used often to describe the quantum measurement problem, where the incoming system becomes entangled with the measuring device. We do not make any claim on what the final state of the particle should be: in many cases, like in a photomultiplier tube, the particle is gone and there is no state. We simply have one incoming state for the particle, prepared over and over exactly the same, and the outgoing state for the measurement device, which is a classical statistical distribution over the different takes.} The question is: how is the statistical distribution coming out of the detector related to the material distribution coming in with the particle?

If the detector does not introduce biases, then the final distribution is only a function of the incoming particle: all unbiased detectors will give the same result. The unstated part also cannot affect the final distribution of outcomes: it is not physically distinguishable. The final statistical distribution of the detector must match the incoming material distribution of the particle. That is if $N$ is the total number of takes and $N_A$ is the number time the detector with aperture in the $A$ region captured the particle, the ration $\frac{N_A}{N}$ will be approximately the amount of material within the aperture $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i$. If we assume an infinite amount of takes, we expect them to coincide: $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i = \lim\limits_{N\rightarrow\infty} \frac{N_A}{N}$. But this is also the frequentist definition of probability. Therefore we have $\int_A \psi^\dagger(q^i) \psi(q^i) dq^i = P(\textrm{measuring particle in } A)$ thus recovering the probabilistic interpretation of the wave function.

We can conceptually extend this picture to the generic case. Suppose we know the outcomes of our measurement are parametrized by a fragment variable $\hat{q}$. We will have a linear functional $\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0} : \Psi \rightarrow \mathbbm{C}$ such that $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2$ returns the amount of material prepared with the particular value $\hat{q}_0$. We can re-express this quantity in terms of the dual and the inner product extended to all Lebesgue integrable functions: $|\Lambda^{\frac{1}{2}}_{\hat{q}=\hat{q}_0}|^2 = |\langle \hat{q}_0 | \psi \rangle|^2 $. This is also the probability that the outcome associated with $\hat{q}_0$ is selected by the measuring device, thus recovering the statistical nature of the quantum predictions.

Note that while this process is of a non-deterministic nature, it is of a specific non-deterministic nature. It requires that no other state can influence the final distribution or the distribution of measurement outcomes would not match the distribution of incoming material. That is: it cannot be understood as an averaging over some unknown external state.

Also note that while we have a way to predict the distribution over final outcomes from the initial state, we have no way of describing what happens in between. When we assume that only the initial state matters, we are saying that we can disregard the particulars of the non-deterministic evolution. So much so that we can substitute it with a deterministic one to compute the probability distribution. In particular, we have no way of predicting which set of final outcomes we will be projecting to, nor when does that actually happen.

While this may seem like an incomplete description, it is as complete as the irreducibility assumption allows. The state is known and determines the final distribution. The unstated part cannot be known, and influences the non-deterministic evolution. If we were able to say something more about the non-deterministic evolution, we would be able to describe an aspect of the unstated part, which we cannot do without violating the assumption. If we cannot learn anything about the unstated part from the final distribution, then we are consistent with our assumption and we are able to use deterministic evolution and projection to determine the final mixed state. In other words: the inability to describe the unstated part must be matched with a corresponding inability to describe its evolution.

We therefore have arrived to all the major underpinning of quantum mechanics and see how they stem from assumption \ref{ass:irreducibility} of irreducibility. We have somewhat a closure, probably not satisfactory to many, for some of the open problems in quantum mechanics. Not because we solved them, we haven't at all, but because we have good reason to have them. The desire to know ``what is really going on" is hindered by assuming that no description can be given below the level of a particle.

On the upside, irreducibility is what it is: an assumption. As with infinitesimal reducibility, it is a reasonable initial assumption assumption. There is hope that, if we abandon it, we can have a more satisfying and complete description. What should be clear, though, that abandoning irreducibility may entail much much more than just finding another equation of motion.
